---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Riccardo Fusaroli"
date: "3/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
p_load(readxl,gtools,tidyverse,dplyr, metafor, lme4)

jinglebells <- read_excel("Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx")
colnames(jingledata)
```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive vocal patterns in schizophrenia? Report how many papers report quantitative estimates, comment on what percentage of the overall studies reviewed they represent (see PRISMA chart) your method to analyze them, the estimated effect size of the difference (mean effect size and standard error) and forest plots representing it. N.B. Only measures of pitch mean and pitch sd are required for the assignment. Feel free to ignore the rest (although pause behavior looks interesting, if you check my article).

2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

## Tips on the process to follow:

- Download the data on all published articles analyzing voice in schizophrenia and the prisma chart as reference of all articles found and reviewed
- Look through the dataset to find out which columns to use, and if there is any additional information written as comments (real world data is always messy!).
    * Hint: PITCH_F0M and PITCH_F0SD group of variables are what you need
    * Hint: Make sure you read the comments in the columns: `pitch_f0_variability`, `frequency`, `Title`,  `ACOUST_ANA_DESCR`, `DESCRIPTION`, and `COMMENTS`
```{r}
#selecting relevant variables
jinglebells <- select(jinglebells,StudyID, SAMPLE_SIZE_HC,SAMPLE_SIZE_SZ,PITCH_F0_HC_M, PITCH_F0_SZ_M, PITCH_F0_HC_SD,PITCH_F0_SZ_SD, PITCH_F0SD_HC_SD,PITCH_F0SD_HC_M,PITCH_F0SD_SZ_M,PITCH_F0SD_SZ_SD, TYPE_OF_TASK) 

#calculating yi and vi
#yi = effect size = mean estimated cohens d within study
#vi = measure of variance = uncertainty of mean

#feature 1
jingledata_mean<-escalc("SMD",n1i = SAMPLE_SIZE_SZ, n2i = SAMPLE_SIZE_HC, m1i =PITCH_F0_SZ_M, m2i =PITCH_F0_HC_M,
          sd1i = PITCH_F0_HC_SD,sd2i=PITCH_F0_SZ_SD,data = jinglebells) 

jingledata_mean <- jingledata_mean[is.na(jingledata_mean$yi) == F,]

#removing study 18, as it was measured in semitones
jingledata_mean <- jingledata_mean[-which(jingledata_mean$StudyID ==18),]

#feature 2
jingledata_SD <- escalc("SMD",n1i = SAMPLE_SIZE_SZ, n2i = SAMPLE_SIZE_HC, 
                     m1i =PITCH_F0SD_SZ_M, m2i =PITCH_F0SD_HC_M,
          sd1i = PITCH_F0SD_SZ_SD,sd2i=PITCH_F0SD_HC_SD,data = jinglebells) 
jingledata_SD <- jingledata_SD[is.na(jingledata_SD$yi) == F,]



#modeling
m1 <- lmer(yi~1+ (1|StudyID),jingledata_mean, weight = 1/vi, EML=F,
     control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore")) #difference between cond

m2<- lmer(yi~1+ (1|StudyID),jingledata_SD, weight = 1/vi, EML=F,
     control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore"))
summary(m1)
summary(m2)

#sample sizes
sum(jingledata_mean$SAMPLE_SIZE_HC)
sum(jingledata_mean$SAMPLE_SIZE_SZ)

sum(jingledata_SD$SAMPLE_SIZE_HC)
sum(jingledata_SD$SAMPLE_SIZE_SZ)


#RMA
pinetrees_mean<- rma(yi,vi,cbind(TYPE_OF_TASK),slab = StudyID, data = jingledata_mean)
forest(pinetrees_mean)

pinetrees_sd<- rma(yi,vi, cbind(TYPE_OF_TASK), slab = StudyID, data = jingledata_SD)
forest(pinetrees_sd)

#loading own data from assingment 3
santa <- read.csv(file.choose())

#extracting relevant variables
santa_wife <- santa %>% 
  group_by(diag) %>% 
  summarise(mean_mean = mean(mean_pitch), 
            sd_sd  = sd(sd_pitch), 
            mean_sd = mean(sd_pitch), 
            sd_mean = mean(mean_pitch))

#making data frames with new data
santa_claus <- data.frame(SAMPLE_SIZE_HC= nlevels(factor(santa$name[santa$diag == 0])),
                          SAMPLE_SIZE_SZ= nlevels(factor(santa$name[santa$diag == 1])),
                          PITCH_F0_HC_M = santa_wife$mean_mean[1],
                          PITCH_F0_SZ_M = santa_wife$mean_mean[2],
                          PITCH_F0_HC_SD = santa_wife$mean_sd[1],
                          PITCH_F0_SZ_SD = santa_wife$mean_sd[2])


santa_claus_2 <- data.frame(SAMPLE_SIZE_HC= nlevels(factor(santa$name[santa$diag == 0])),
                          SAMPLE_SIZE_SZ= nlevels(factor(santa$name[santa$diag == 1])),
                          PITCH_F0SD_HC_M = santa_wife$sd_mean[1],
                          PITCH_F0SD_SZ_M = santa_wife$sd_mean[2],
                          PITCH_F0SD_HC_SD = santa_wife$sd_sd[1],
                          PITCH_F0SD_SZ_SD = santa_wife$sd_sd[2])

#calculating yi and vi for new data
jinglesanta<-escalc("SMD",n1i = SAMPLE_SIZE_SZ, n2i = SAMPLE_SIZE_HC, 
                    m1i =PITCH_F0_SZ_M, m2i =PITCH_F0_HC_M,
          sd1i = PITCH_F0_SZ_SD,sd2i=PITCH_F0_HC_SD,data = santa_claus) 

jinglesanta_2<-escalc("SMD",n1i = SAMPLE_SIZE_SZ, n2i = SAMPLE_SIZE_HC, 
                      m1i =PITCH_F0SD_SZ_M, m2i =PITCH_F0SD_HC_M,
          sd1i = PITCH_F0SD_SZ_SD,sd2i=PITCH_F0SD_HC_SD,data = santa_claus_2)

#binding new and old dataframes
christmes<- smartbind(jingledata_mean, jinglesanta)
christmes$StudyID[is.na(christmes$StudyID) == T] <- 100
christmes_2<- smartbind(jingledata_SD, jinglesanta_2)
christmes_2$StudyID[is.na(christmes_2$StudyID) == T] <- 101

sum(christmes$SAMPLE_SIZE_HC)
sum(christmes$SAMPLE_SIZE_SZ)

sum(christmes_2$SAMPLE_SIZE_HC)
sum(christmes_2$SAMPLE_SIZE_SZ)

#RMA
elf<- rma(yi,vi, cbind(TYPE_OF_TASK), slab = StudyID, data = christmes)
forest(elf)

elf_2<- rma(yi,vi, cbind(TYPE_OF_TASK), slab = StudyID, data = christmes_2)
forest(elf_2)
```

```{r}
#Summary of rma
summary(pinetrees_mean)
summary(pinetrees_sd)

summary(elf)
summary(elf_2)
#Testing for influential studies
inf<- influence(elf)
plot(inf)
inf <- influence(elf)
plot(inf)

#kendalls tau
ranktest(elf)
ranktest(elf_2)


#funnel plots
funnel(pinetrees_mean, main = "Pitch Mean", xlab = "Standardized Mean Difference")
funnel(pinetrees_sd, main = "Pitch variance", xlab = "Standardized Mean Difference")
funnel(elf, main = "Pitch Mean incl. ass3 data", xlab = "Standardized Mean Difference")
funnel(elf_2_nosix, main = "Pitch Variance incl. ass3 data", xlab = "Standardized Mean Difference")

#extracting study 15 because it was too influential
christmes_2_nosix <- christmes_2[-6,]

#rerunning analysis
elf_2_nosix<- rma(yi,vi, cbind(TYPE_OF_TASK), slab = StudyID, 
                  data = christmes_2_nosix)

sum(christmes_2_nosix$SAMPLE_SIZE_HC)
sum(christmes_2_nosix$SAMPLE_SIZE_SZ)

forest(elf_2_nosix)

ranktest(elf_2_nosix)
```

- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))
 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

- Now look at the output of rma() and check tau and I2
